---
title: "Cleaning and exploratory analysis"
author: "Juan Berrios"
date: "6/18/2022"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: show
    df_print: paged
---

```{r setup, echo=F}
knitr::opts_chunk$set(include=TRUE, echo=TRUE, comment=NA)
```

```{css, echo=F}
pre {
  max-height: 300px;
  overflow-y: auto;
}
```

# Introduction

- The following is code written for an analysis of diminutive formation in L2 Spanish. The data come from [Corpus Escrito del Español como L2 (CEDEL2)](http://cedel2.learnercorpora.com/). It follows a first notebook concentrated on locating and tagging tokens for later analyses. 

# Preparation

- Loading libraries:

```{r message=FALSE,warning=FALSE}
library(tidyverse)
library(magrittr)
```

- Loading data:

```{r}
#Note that files is tab separated.

read.csv("./data/diminutives.csv") -> df
read.csv("./data/texts_processed.csv") -> texts
```

- Checking that everything was loaded properly:

```{r}
df %>%
  colnames()
```
```{r}
df %>%
  head()
```

```{r}
df %>%
  dim()
```

- Looks like there are 6,696 tokens after a first attempt. Many of these, however, are not diminutives but rather items that happen to end in the same combination of segments. A first attempt to remove non-diminutives is to filter out highly frequent items, which, by virtue of their frequency tend to be more semantically independent (lexicalized). We'll use lemmas to be more efficient (collapsing masc/fem and sing/plural categories):

```{r}
#Simple frequency distribution for tokens (will sort out some non-diminutives). This list includes those that appear at least twice in the data frame. 

frequent_lemmas <- df %>%
  count(Lemma, sort = TRUE) %>%
  filter(n > 2) %>%
  rownames_to_column()
```


```{r}
frequent_lemmas
```


```{r}
#Making  lists based on results

indices <- c(2,4,11,15,20,26,32,33,36,45,50,52,56,58,63,64,69,71,76,82,83,85,87,89,96,102,106,107,109,110,111,113:115,118,120,122,123,126,127,132:135,138,139,142,144,147,149:151,157,158,162,163,165,173:176,178,180:182,187,199:201,204,208,210)

diminutives <- frequent_lemmas$Lemma[indices]
non_diminutives <- frequent_lemmas$Lemma[-indices]
```

- List of diminutives kept:

```{r}
print(diminutives)
```

- List of non-diminutives removed:

```{r}
print(non_diminutives)
```

- Remove non-diminutives from data frame:

```{r}
df <- df %>% 
    filter(!Lemma %in% non_diminutives)

#Dimensions
print("There are this many data points as of now:")
dim(df)[1]
print("There are this many unique essays which tokens appear in:")
length(unique(df$Filename))
```
- As a last step we can examine hapax legomena (those that appear only once):

```{r}
hapax <- df %>%
  count(Lemma, sort = TRUE) %>%
  filter(n == 1) %>%
  rownames_to_column()
```

- As can be seen below, many appear to be lemmatization/tokenization errors, we'll fix them as appropriate:

```{r}
hapax
```

```{r}
#Making  lists based on results

indices <- c(3,4,7,10,12,15,19,20,21,23,25,30,31,33,36,40,42,44,45,52,53,54,56:60,62:64,66:68,70,71:76,79,80:83,85,86,89,90,91,92,94:99,101:103,105:107,112,116,118,119,121:122,124,126:130,133,135:137,141,143,144,146,150,151,153:155,158,168,171:173,175,176,181:183,188:190,192:197,206,210,211:217,220:222,224:234,236,239,240,242,244:246,248:250,252:256,274,276,277,278,280:283,285,286,287,288,290,292,294,296,297,298:300,302,303,306,307,308,310,311,312,315,317,320,321,323,325,327,329,332,334,335,336,337,338,340,341,342,343,346:348,353,354:359,361,363,364,366,368:370,371:373,376,379,380,381,385,387:389,391,392,393,394,396,401)

diminutives <- hapax$Lemma[indices]
non_diminutives <- hapax$Lemma[-indices]
```

- Remove non-diminutives from data frame:

```{r}
df <- df %>% 
    filter(!Lemma %in% non_diminutives)

#Dimensions
print("There are this many data points as of now:")
dim(df)[1]
print("There are this many unique essays which tokens appear in:")
length(unique(df$Filename))
```

```{r}
print("There are this many tokens by diminutive used:")
count(df, Diminutive)
```


# Brief frequency analysis

- To begin examining pattern, we can do a brief frequency analysis:

```{r warning=FALSE}
#Loading stop words

tibble(Token = quanteda::stopwords("es"),Lexicon = "custom") -> stop_words
```

- For all tokens accross the data set:

```{r}
#Simple frequency distribution for tokens

texts %>%
  filter(POS != 'SPACE' & POS != 'PUNCT') %>% #remove punctuation
  anti_join(stop_words %>% filter(Lexicon=="custom")) %>% #remove stop words
  count(Token, sort = TRUE) %>%
  head(25) #Top 25
```

- For the diminutive subset only:

```{r}
#Simple frequency distribution for tokens

df %>%
  count(Token, sort = TRUE) %>%
  head(25) #Top 25
```

- Plotting:

```{r echo=FALSE}

#Function to set size 

psize <- function(width_choice, height_choice) {
        options(repr.plot.width=width_choice, repr.plot.height=height_choice)
        }
```

- All tokens:

```{r echo=FALSE}
psize(12, 8)

texts %>%
  filter(POS != 'SPACE' & POS != 'PUNCT') %>% #remove punctuation
  anti_join(stop_words %>% filter(Lexicon=="custom")) %>% #remove stop words
  count(Token, sort = TRUE) %>%
  head(25) %>%
  mutate(Token = reorder(Token, n)) %>%
  ggplot(aes(n, Token, fill=Token)) +
  geom_col(color="black") +
  labs(title="Frequency Distribution of Tokens",x='Counts',y = NULL,fill='Token') +
  theme_classic()+
  theme(text = element_text(size=20))

ggsave("figures/frequency_distribution.png",width = 12, height = 8,dpi=300)
```

- Diminutives only:

```{r echo=FALSE}
psize(12, 8)

df %>%
    count(Token, sort = TRUE) %>%
    head(25) %>%
    mutate(Token = reorder(Token, n)) %>%
    ggplot(aes(n, Token, fill=Token)) +
    geom_col(color="black") +
    labs(title="Frequency Distribution of Tokens",x='Counts',y = NULL,fill='Token') +
    theme_classic()+
    theme(text = element_text(size=20))

ggsave("figures/frequency_distribution_diminutives.png",width = 12, height = 8,dpi=300)
```

- Cross linguistic analyses:

```{r}
df %>%
    count(L1, sort = TRUE) %>%
    mutate(L1 = reorder(L1, n)) %>%
    ggplot(aes(n, L1, fill=L1)) +
    geom_col(color="black") +
    labs(title="Tokens by L1",x='Counts',y = NULL,fill='L1') +
    theme_classic()+
    theme(text = element_text(size=20))
```

- Tokens by L1:

```{r}
df %>% 
    count(L1, sort = TRUE) 
```

- Tasks by L1:

```{r}
df %>%
  group_by(Task.title, L1) %>%
  tally() %>%
  spread(L1, n)
```

- Saving the different objects for later analyses:

```{r}
#Data frames

write.csv(df,"data/diminutives_cleaned.csv",row.names = FALSE)

#Lists
write.table(diminutives, "data/diminutives.txt",sep="\t", row.names=FALSE, col.names = FALSE)
write.table(non_diminutives, "data/non_diminutives.txt",sep="\t", row.names=FALSE,col.names = FALSE)
```


# Session info

```{r}
sessionInfo()
```