---
title: "Cleaning and exploratory analysis"
author: "Juan Berrios"
date: "6/30/2022"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
    number_sections: yes
    code_folding: show
    df_print: paged
---

```{r setup, echo=F}
knitr::opts_chunk$set(include=TRUE, echo=TRUE, comment=NA)
```

```{css, echo=F}
pre {
  max-height: 300px;
  overflow-y: auto;
}
```

# Introduction

- The following is code written for an analysis of diminutive formation in L2 Spanish. The data come from [Corpus Escrito del Español como L2 (CEDEL2)](http://cedel2.learnercorpora.com/). It follows a first notebook concentrated on locating and tagging tokens for later analyses, and a cleaning notebook to removed unneeded tokens and fix inaccurate tags. 

# Preparation

- Loading libraries:

```{r message=FALSE,warning=FALSE}
library(tidyverse)
library(magrittr)
library(lme4)
```

- Loading data:

```{r}
#Note that files are tab separated.

read.csv("./data/diminutives_analysis.csv",fileEncoding = "UTF-8-BOM") -> df
read.csv("./data/texts_processed.csv") -> texts
read.csv("./data/texts.csv",sep='\t') -> texts_original_L2
read.csv("./data/native/texts.csv",sep='\t') -> texts_original_native
```

- Checking that everything was loaded properly for diminutives data frame:

```{r}
df %>%
  colnames()
```
```{r}
df %>%
  head()
```

- Likewise for texts dataframe:

```{r}
texts %>%
  colnames()
```

```{r}
texts %>%
  dim()
```

- And lastly for the original texts, which we'll merge first:

```{r}
texts_original_L2 <- texts_original_L2 %>%
  select(Subcorpus,Filename,L1,Task.title,Text)

texts_original_native <- texts_original_native  %>%
  select(Subcorpus,Filename,L1,Task.title,Text)

texts_original <- texts_original_L2 %>%
                    full_join(texts_original_native)

#Save memory
rm(texts_original_L2)
rm(texts_original_native)
```

- Checking:

```{r}
texts_original %>%
  colnames()
```

```{r}
texts_original %>%
  dim()
```
- Removing L2s (keeping English, German, Portuguese, Italian, Russian, Spanish) and tasks (keeping only narratives) that we will not consider.

```{r}
L1s = c("English","German","Portuguese","Italian","Russian","Spanish")
tasks = c("3. Film","4. Last year holidays","6. Recent trip","7. An experience", "13. Frog","14. Chaplin")
```

```{r}
df <- df %>% 
  filter(Task.title %in% tasks & L1 %in% L1s)
```

```{r}
texts <- texts %>% 
  filter(Task.title %in% tasks & L1 %in% L1s)
```

```{r}
texts_original <- texts_original %>% 
  filter(Task.title %in% tasks & L1 %in% L1s)
```

- Adding a column for accuracy in L2 dataset:


```{r}
df <- df %>% 
  mutate(Accuracy = case_when(Token.corrected == Token.checked ~ "Accurate",
                        Token.corrected != Token.checked  ~ "Innacurate"))
```

# Exploring

```{r}
print("There are this many tokens as of now:")
dim(df)[1]
print("There are this many types as of now:")
length(unique(df$Token.corrected))
print("There are this many types (spellecheked) as of now:")
length(unique(df$Token.checked))
print("There are this many lemmas as of now:")
length(unique(df$Lemma.corrected))
print("There are this many unique essays which tokens appear in:")
length(unique(df$Filename))
```


```{r}
print("There are this many tokens by diminutive used:")
count(df, Diminutive)
```


# Brief frequency analysis

- To begin examining pattern, we can do a brief frequency analysis:

```{r warning=FALSE}
#Loading stop words

tibble(Token = quanteda::stopwords("es"),Lexicon = "custom") -> stop_words
```

- For all tokens accross the data set:

```{r}
#Simple frequency distribution for tokens

texts %>%
  filter(POS != 'SPACE' & POS != 'PUNCT') %>% #remove punctuation
  anti_join(stop_words %>% filter(Lexicon=="custom")) %>% #remove stop words
  count(Token, sort = TRUE) %>%
  head(25) #Top 25
```

- For the diminutive subset only:

```{r}
#Simple frequency distribution for tokens

df %>%
  count(Token, sort = TRUE) %>%
  head(25) #Top 25
```

- Plotting:

```{r echo=FALSE}

#Function to set size 

psize <- function(width_choice, height_choice) {
        options(repr.plot.width=width_choice, repr.plot.height=height_choice)
        }
```

- All tokens:

```{r echo=FALSE}
psize(12, 8)

texts %>%
  filter(POS != 'SPACE' & POS != 'PUNCT') %>% #remove punctuation
  anti_join(stop_words %>% filter(Lexicon=="custom")) %>% #remove stop words
  count(Token, sort = TRUE) %>%
  head(25) %>%
  mutate(Token = reorder(Token, n)) %>%
  ggplot(aes(n, Token, fill=Token)) +
  geom_col(color="black") +
  labs(title="Frequency Distribution of Tokens",x='Counts',y = NULL,fill='Token') +
  theme_classic()+
  theme(text = element_text(size=20))

ggsave("figures/frequency_distribution.png",width = 12, height = 8,dpi=300)
```

- Diminutives only:

```{r echo=FALSE}
psize(12, 8)

df %>%
    count(Token, sort = TRUE) %>%
    head(25) %>%
    mutate(Token = reorder(Token, n)) %>%
    ggplot(aes(n, Token, fill=Token)) +
    geom_col(color="black") +
    labs(title="Frequency Distribution of Tokens",x='Counts',y = NULL,fill='Token') +
    theme_classic()+
    theme(text = element_text(size=20))

ggsave("figures/frequency_distribution_diminutives.png",width = 12, height = 8,dpi=300)
```

- Cross linguistic analyses:

```{r}
df %>%
    count(L1, sort = TRUE) %>%
    mutate(L1 = reorder(L1, n)) %>%
    ggplot(aes(n, L1, fill=L1)) +
    geom_col(color="black") +
    labs(title="Tokens by L1",x='Counts',y = NULL,fill='L1') +
    theme_classic()+
    theme(text = element_text(size=20))
```

- Tokens by L1:

```{r}
df %>% 
    count(L1, sort = TRUE) 
```

- Tasks by L1:

```{r}
df %>%
  group_by(Task.title, L1) %>%
  tally() %>%
  spread(L1, n)
```

- Error rates:

```{r}
count(filter(df,L1 != 'Spanish'),Accuracy)
```
- Part of speech:

```{r}
count(df,POS)
```
## Relative frequency

- The figures above, however, can be deceiving as the subsets (by L1) are not equal in size. We'll now do a relative frequency analysis by L1 and by individual participant. We'll start with L1s:

```{r}
original <- texts_original %>%
  group_by(L1) %>%
  summarize(n=n()) %>%
  rename(Essays = n)

processed <- texts %>%
  distinct(Filename, .keep_all = TRUE) %>%
  group_by(L1) %>%
  summarize(n=n()) %>%
  rename(Essays.with.diminutive = n)
```

```{r}
original %>%
  full_join(processed,by="L1") %>%
  mutate(Percent.used = Essays.with.diminutive * 100 / Essays)
```

- Number of words out of total words that are diminutives:

```{r}
total_words <- texts %>%
  group_by(L1) %>%
  filter(POS != 'SPACE' & POS != 'PUNCT') %>%
  summarize(n=n()) %>%
  rename(Words = n)

diminutives <- df %>%
  group_by(L1) %>%
  summarize(n=n()) %>%
  rename(Diminutives = n)
```

```{r}
total_words %>%
  full_join(diminutives,by="L1") %>%
  mutate(Percent.used = Diminutives * 100 / Words)
```
- By learner:

```{r}
total_words <- texts %>%
  group_by(Filename) %>%
  filter(POS != 'SPACE' & POS != 'PUNCT') %>%
  summarize(n=n()) %>%
  rename(Words = n)

diminutives <- df %>%
  group_by(Filename) %>%
  summarize(n=n()) %>%
  rename(Diminutives = n)
```

```{r}
percent_used <- total_words %>%
  full_join(diminutives,by="Filename") %>%
  drop_na() %>% 
  mutate(Percent.used = Diminutives * 100 / Words) %T>%  print()

#NOTE: Do this same analysis later by learner, so each can have a percentage of diminutives used as an independent variable
```

- Attaching percetages to main data frame so that they can be used as a variable:

```{r}
df <- percent_used %>%
  select(Filename,Percent.used) %>%
  full_join(df,by="Filename") %T>% print()
```

# Linear modeling:

## Accuracy

```{r}
#L1 (including Spanish)
glm(as.factor(Accuracy) ~ L1, family=binomial,df) %>% summary()
```

```{r}
#L1 (excluding Spanish)
glm(as.factor(Accuracy) ~ L1, family=binomial,filter(df, L1 != 'Spanish')) %>% summary()
```
```{r}
#Proficiency score (excluding Spanish)
glm(as.factor(Accuracy) ~ Proficiency.score, family=binomial(link = "logit"),filter(df, L1 != 'Spanish')) %>% summary()
```
```{r}
#Proficiency range (excluding Spanish)
glm(as.factor(Accuracy) ~ Proficiency.range, family=binomial(link = "logit"),filter(df, L1 != 'Spanish')) %>% summary()
```
```{r}
#POS (including Spanish)
glm(as.factor(Accuracy) ~ POS, family=binomial(link = "logit"),df) %>% summary()
```

```{r}
#POS (excluding Spanish)
glm(as.factor(Accuracy) ~ POS, family=binomial(link = "logit"),filter(df, L1 != 'Spanish')) %>% summary()
```
```{r}
#POS (excluding adverbs)
glm(as.factor(Accuracy) ~ POS, family=binomial(link = "logit"),filter(df, POS != 'ADV')) %>% summary()
```
```{r}
#POS (excluding adverbs and Spanish L1)
glm(as.factor(Accuracy) ~ POS, family=binomial(link = "logit"),filter(df, POS != 'ADV' & L1 != 'Spanish')) %>% summary()
```
## Percent used

```{r}
#L1 (including Spanish)
lm(Percent.used ~ L1, df) %>% summary()
```

```{r}
#L1 (excluding Spanish)
lm(Percent.used ~ L1, filter(df, L1 != 'Spanish')) %>% summary()
```

```{r}
#Proficiency score (excluding Spanish)
lm(Percent.used ~ Proficiency.score,filter(df, L1 != 'Spanish')) %>% summary()
```
```{r}
#Proficiency range (excluding Spanish)
lm(Percent.used ~ Proficiency.range, filter(df, L1 != 'Spanish')) %>% summary()
```

```{r}
#POS (including Spanish)
lm(Percent.used ~ POS, df) %>% summary()
```

```{r}
#POS (excluding Spanish)
lm(Percent.used ~ POS, filter(df, L1 != 'Spanish')) %>% summary()
```

```{r}
#POS (excluding adverbs)
lm(Percent.used ~ POS, filter(df, POS != 'ADV')) %>% summary()
```

```{r}
#POS (excluding adverbs and Spanish L1)
lm(Percent.used ~ POS, filter(df, POS != 'ADV' & L1 != 'Spanish')) %>% summary()
```
```{r}
write.csv(df, file='data/diminutives_regressions.csv')
```

# Session info

```{r}
sessionInfo()
```